\acknowledgments{% \vskip2pc
  % {\narrower\noindent
It takes a village, and without my village, this thesis wouldn't have happened.
My parents, Amy and Brian, always encouraged me to pursue my interests,
  and they put up with seven-year-old me's endless appetite for algebra problems.
I've always looked up to my big brother Sam, even though he's two years younger,
  and to his family, Alexis and Charlie.

Zach: you've been everything I needed in an advisor. I couldn't have done it without you.
And a special thank you for putting so much time into building the lab into what it is today.
To my reading committee members, Dan and Tom, you are both such great role models for how to be an academic.
I learned a ton from watching you work.
To my GSR and friend, Geoffrey, my time in your choir was
some of the only consistent (and much needed) structure to my schedule those first few years in grad school,
and really helped me feel connected to UW and Seattle.

To all the members of the PLSE lab, what an incredible space to learn and hang out!
To Doug, I learned so much from you (and your dad). Thanks for teaching me how to impersonate a systems person.
Thanks to Colin, Konstantin, Pavel, Eric, Stuart, Calvin, Talia, John, Chandra, Max, and Remy for teaching me so much about PL
while disguising it as shooting the breeze.
Also a big thanks to all the incredible UW undergraduates I worked with over the years,
  Steve, Ryan, Pooja, Justin, Miranda, David, Ethan, and Taylor.
You all contributed so much to our projects!

One of the best parts about getting a PhD is becoming a member of the international research community.
I've had the pleasure of meeting many wonderful people and starting great collaborations with several of them.
I met Ilya at the PLDI 2015 banquet, and it wasn't long before we were talking about the basic ideas of \disel.
I met Karl at the ASE 2015 banquet, and it wasn't long before we were talking about Verdi and sensor networks.
I met Oded at PLDI 2016, and it wasn't long before we were working on yet another proof of Raft's correctness.
Oded introduced me to Mooly, Yotam, Sharon, and Marcelo, as well as to Ken and Giuliano.
Everyone in this entourage has helped me realize that I am a collaborator at heart.
Here's to many more projects together!

I had a beautiful time at my summer internship at Microsoft Research with Jay Lorch and Rustan Leino.
To this day, one of my favorite things to do is to make Jay laugh.
Rustan taught me pretty much everything I know about building verification tools,
especially the insight that compiler engineering principles apply to them.
That summer, I also had the chance to get to know Shaz Qadeer,
who taught me how to think about refinement proofs and concurrency.

Finishing this thesis was not an easy journey for me.
I left UW having defended but not submitted the document in June 2019,
and at that point I was feeling pretty ground down and not able to make much progress.
I consider it a small miracle that the document will finally be submitted in the summer of 2021.
I found a very friendly and welcoming home in the intervening years at Certora,
with Mooly and Shelly and the rest of the team.
Big thanks to them for hiring me, and for keeping me around part time when I tried to quit.
While working at Certora, Daniel Ricketts and Calvin at Oracle were excellent lunch buddies.

I wouldn't have been able to get back to making progress on the thesis without Zach's gentle guidance and project management.
``Five minutes a day'' was life changing for me a year ago,
and I'm so happy to have seen it grow into a full community, first with Chandra and then with \texttt{\#accountability-coop}.
Finally, thanks to Race Condition Running (miles, not pace!), the choirs of Saint Mark's cathedral,
and the species \emph{Canis familiaris} for supporting my mental health through this process.

  % \par}
}
