\documentclass[11pt, proquest]{uwthesis}[2016/11/22]

\usepackage{amsmath}
\usepackage{setspace}
\usepackage{mathpartir}
\usepackage[usenames,dvipsnames]{color}
\usepackage{alltt}
\usepackage{textcomp}
\usepackage{amscd}
\usepackage{amsfonts}
\usepackage{varwidth}
\usepackage{listings}
\usepackage{float}
\usepackage[noend]{algorithmic}
\usepackage{mathrsfs}
\usepackage{stmaryrd}
\usepackage{verbatim}
\usepackage{lipsum}
\newcounter{tags}
\usepackage{enumitem}
\usepackage{fancyvrb}
\fvset{fontsize=\scriptsize}
\renewcommand{\ttdefault}{txtt} % provide bold tt in alltt

\lstdefinelanguage{mypyvy}
{morekeywords={mutable, immutable, constant, relation, function, 
               init, transition, modifies, new, safety, invariant,
               sort, forall},
sensitive=false,
morecomment=[l]{\#},
basicstyle=\small\ttfamily,
keywordstyle=\small\ttfamily\bfseries,
}

%\usepackage{libertine}
\usepackage[T1]{fontenc}

\usepackage[numbers]{natbib}

\usepackage{graphicx}
\usepackage{pifont}
\usepackage{xspace}
\usepackage{relsize}
\PassOptionsToPackage{hyphens}{url}\usepackage[colorlinks=true,allcolors=black,breaklinks,draft=false]{hyperref}
\usepackage{amssymb}
\usepackage{wasysym}
\usepackage{amsthm}
\usepackage[noabbrev,capitalise]{cleveref}
\theoremstyle{definition}
\newtheorem{theorem}{Theorem}[section]
\newtheorem{definition}[theorem]{Definition}
\newtheorem{lemma}[theorem]{Lemma}
\newtheorem{proposition}[theorem]{Proposition}
\newtheorem{corollary}[theorem]{Corollary}

\newcounter{eventnum}\setcounter{eventnum}{0}
\renewcommand{\theeventnum}{\Alph{eventnum}}
\newcommand{\mkevent}[1]{\refstepcounter{eventnum}\label{#1}}

\Crefname{eventnum}{event}{events}
\Crefname{thm}{Theorem}{Theorems}
\Crefname{lem}{Lemma}{Lemmas}
\Crefname{chapter}{Chapter}{Chapters}

\usepackage{tikz}
\usetikzlibrary{calc}
\usetikzlibrary{shapes.callouts}
\usetikzlibrary{arrows.meta}
\usetikzlibrary{decorations.pathreplacing}


\usepackage{multirow}
\usepackage{booktabs}

\setcounter{tocdepth}{1}

\input{macros}

\begin{document}

\prelimpages

\Title{Compositional and Automated Verification of Distributed Systems}
\Author{James R. Wilcox}
\Year{2021}
\Program{\pgascse}

\Chair{Zachary Tatlock}{Professor}{\pgascse}
\Signature{Dan Grossman}
\Signature{Thomas Anderson}

\titlepage

\copyrightpage

\setcounter{page}{-1} \abstract{%
  Distributed systems provide the backbone for modern computer
  systems, from cloud computing to air-traffic control.  These complex
  systems execute concurrently in unreliable environments and are
  expected to tolerate various faults.  Such environments are
  notoriously difficult to adequately model with testing, but because
  of the critical importance of these systems, it is essential that
  they are correct.  It thus makes sense to turn to more rigorous
  methods of ensuring correctness, such as formal verification.
  Applying formal methods is not a panacea, however, due to the
  complexity of the systems involved.  It is not uncommon, \eg, for a
  distributed file system to coordinate thousands of machines using a
  combination of several different protocols to ensure consistency,
  fault tolerance, and high performance.  Verifying such a
  system requires breaking the problem down into individually
  verifiable parts, and leveraging automation whenever possible.

  This dissertation describes programming languages techniques for
  verifying distributed systems
  compositionally and automatically.  First, we present \Verdi, a
  framework for verifying distributed systems that reasons about fault tolerance mechanisms
  as transformers between fault models. Second, we detail \disel, a
  concurrent separation logic for distributed systems whose key
  insight is to treat the network as analogous to the heap in
  sequential programming.  Finally, we report on \mypyvy, a
  domain-specific language for symbolic transition systems
  in first-order logic, which supports a variety of automated
  reasoning tools to analyze systems.
}

\tableofcontents
%\listoffigures
%\listoftables  % I have no tables

\acknowledgments{% \vskip2pc
  % {\narrower\noindent
  Acks.
  % \par}
}

%\dedication{\begin{center}dedication\end{center}}

\textpages

\input{intro}
\input{verdi}
\input{disel}
\input{mypyvy}
\input{conclusion}

%\vita{
%  Vita
%}

\bibliographystyle{plainnat}
\bibliography{refs}


\end{document}
